---
title: "elliet85_HW_10"
author: "Ellie Terry"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  bookdown::html_document2:
    theme: cosmo
    toc: true
    toc_float: true
    bibliography: hw10.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(pander)
library(readxl)
library(kableExtra)
library(captioner)

```

```{r}
table_nums <- captioner(prefix = "Table")
figure_nums <- captioner(prefix = "Figure")
```

```{r import_data}
births <- read.table("ISLbirth.txt", sep=",", header = T) 

births <- births %>%
  set_names(~ str_to_lower(.) %>%
              str_replace_all(" ", "_")) %>% 
 filter(year %in% c(1960:2018)) %>% 
 select(year, births, sex)

deaths <- read.table("ISLdeath.txt", sep=",", header = T) 
deaths <- deaths %>%
  set_names(~ str_to_lower(.) %>%
              str_replace_all(" ", "_")) %>% 
  filter(year %in% c(1960:2018)) %>% 
  select(year,age, deaths, sex) 
  
deaths2 <-deaths %>% 
  group_by(year,sex) %>% 
   summarise(deaths=sum(deaths))

#gdp <- gdp%>% 
#  filter(year %in% c(1960:2018)) %>% 
#  mutate(year = as.integer(year))
#gdp <- read_excel("Iceland_gdp.xls") 

```

#Research question

In this paper I use Data from the Human Mortality Database to examine birth and death trends in Iceland

```{r bind_data}
iceland <- as.data.frame(left_join(births, deaths2))
```

#Analysis

The figure below shows how births and deaths have changed over time. Notably there is an upward trend in deaths Iceland with a spike in 2016. 
```{r, fig.cap= "Births and Deaths in Iceland: 1960-2018"}
ggplot(data = iceland,
 aes(x = year)) +
  geom_line(aes(y = births, color="Births"), size = 1) +
  geom_line(aes(y = deaths, color="Deaths"), size = 1) +
  ylab("Number of Births and Deaths") +
  xlab("Year") +
  scale_color_manual("",
                     breaks = c("Births", "Deaths"),
                     values = c("goldenrod", "skyblue3")) +
  ggtitle("Births and Deaths in Iceland, 1960-2018") +
  theme(legend.position = "bottom")
 
```

What was the source of the increase in deaths in 2016? The table below shows deaths by age group in 2016 versus 2017.
```{r}

table2 <- deaths %>%
  filter(year %in% c(2016,2017)) %>% 
  group_by(year,age) %>% 
  summarise(`Total Deaths` = sum(deaths)) %>% 
  arrange(age,year)
  
kable(table2, caption = "Deaths by Age in 2016 vs 2017") %>% 
  kable_styling(full_width = T, position = "center")
```

#Results
The Results show that the source of the spike in 2016 occurred in nearly all age groups. 

#References
---
nocite: '@*'
---


